---
title: "Homework 1 - Monte Carlo Methods"
author: "Zhuohui(Jeffrey) Liang ZL2974"

output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

# Problem 1
The standard Laplace distribution has density $f(x) = 0.5e^{-|x|}, x\in
(-\infty, \infty)$. Please provide an algorithm that uses the inverse transformation method to generate a random sample from this distribution. Use the $U(0,1)$ random number generator in  \em{\bf{R}}, write a  \em{\bf{R}}-function to implement the algorithm. Use visualization tools to validate your algorithm (i.e., illustrate whether the random numbers generated from your function truely follows the standard Laplace distribution.)

# Answer: your answer starts here...

$$F(X) =
\begin{cases}
\int_{-\infty}^x{0.5e^{x}}dx = 0.5e^x & \text{if x <0;}\\
0.5 + \int_{0}^x{0.5e^{-x}}dx = 1- 0.5e^{-x} & \text{if}~x \ge 0\\
\end{cases}$$

s.t. 
$$F^{-1}(u) =
\begin{cases}
\log{2u} & \text{if u <0.5;}\\
-\log{(2-2u)} & u \in [0.5,1]
\end{cases}$$

```{r p1}
rlaplace = 
  function(number_of_randomize){
    unif_vec = runif(number_of_randomize)
    laplace_vec = 
      (unif_vec < 0.5) * log(2*unif_vec) -
      between(unif_vec,0.5,1) * log(2 - 2*unif_vec)
    return(laplace_vec)
  }

rlaplace(10000) %>% 
  as_tibble() %>% 
  ggplot(aes(x = value)) +
  geom_density(adjust = 2)

```
$\blacksquare$

#Problem 2

Use the inverse transformation method to  derive an algorithm for generating a Pareto random number with $U\sim U(0,1)$, where the Pareto random number has a probability density function 
     $$f(x; \alpha, \gamma)=\frac{\gamma\alpha^{\gamma}}{x^{\gamma+1}} I\{x\ge \alpha\}$$ with two parameters $\alpha >0$ and $\gamma>0$. Use visualization tools to validate your algorithm (i.e., illustrate whether the random numbers generated from your function truely follows the target distribution.)

\begin{eqnarray*}
F(x; \alpha, \gamma) & = & \gamma\alpha^\gamma\int_\alpha^x{s^{-(\gamma+1)}ds}\\
& = & 1 - \alpha^\gamma * x^{-\gamma}
\end{eqnarray*}

\begin{eqnarray*}
F^{-1}(u; \alpha, \gamma) & = & (\frac{1-u}{\alpha^\gamma})^{-\frac{1}{\gamma}}\\
&=& \alpha *(1-u)^{-\frac{1}{\gamma}}
\end{eqnarray*}

```{r}
rpareto =
  function(number_of_randomize, alpha, gamma) {
    
    u = runif(number_of_randomize)
    pareto =
      alpha * (1-u)^(-1/gamma)
    
    return(pareto)
  }

set.seed(123123)
a = 2
rpareto(10000,a,2) %>%
  as_tibble() %>%
  ggplot(aes(x = value)) +
  geom_density(adjust = 2) +
  geom_vline(aes(xintercept = a))
```

#Problem 3

Construct an algorithm for using the acceptance/rejection method to generate 100 pseudorandom
 variable from the pdf
 $$f(x) = \frac{2}{\pi \beta^2} \sqrt{\beta^2-x^2}, \,\, -\beta \le x \le \beta.$$
The simplest choice for $g(x)$ is the $U(-\beta, \beta)$ distribution but other choices are possible as well. Use visualization tools to validate your algorithm (i.e., illustrate whether the random numbers generated from your function truely follows the target distribution.)

# Answer:  your answer starts here...
```{r }
#Your R codes/functions 
```





