---
title: "Homework 01-25-2021"
author: "Jeffrey Liang ZL2974"

output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```


# Problem 1

Develop two Monte Carlo methods for the estimation of $\theta=\int_0^1 e^{x^2} dx$ and implement in \em{\bf{R}} .

# Answer:  your answer starts here...

Method 1

Introduce Let $Y\sim U(0,1)$
$$\theta = \int_0^1{e^{y^2}*1dx}=E[e^{y^2}]$$
```{r method_1}
set.seed(123123)
y = runif(1000, 0, 1)

gfun = function(y) {
  return((exp(y ^ 2)))
}

gx =
  gfun(y)

theta_estimate =
  list(theta = mean(gx),
       var_theta = var(gx))

integrate(function(x)
  exp(x ^ 2), 0, 1)
```

Method 2
```{r}
plot.function(gfun,0,1)
```

looking at the function, which is similar to a exponetial funciton, thus 
introducing $f(x) = e^x$ as control variate, we have:

$$\theta = \int_0^1{\beta*e^x + (e^{x^2}-\beta*e^x)dx} = \beta*e^x|^1_0+E[(e^{x^2}-\beta*e^x)]$$
```{r method2}
ffun = function(y){
  return(exp(y))
}

fx = ffun(y)

b = lm(gx~fx)$coef[2]

theta_estimate_2 =
  list(theta = mean(b*(exp(1)-1)+(gx-b*fx)),
       theta2 = sum(lm(gx~fx)$coef*c(1,mean(fx))),
       var_theta =var((b*(exp(1)-1)+(gx-b*fx)))
       )

theta_estimate_2
  
```



# Problem 2
Show that in estimating $\theta=E\{\sqrt{1-U^2}\}$ it is better to use $U^2$ rather than $U$ as the control variate, where $U\sim U(0,1)$.  To do this, use simulation to approximate the necessary covariances. In addition, implement your algorithms in  \em{\bf{R}}.

# Answer:  your answer starts here...
```{r q2}
set.seed(123123)

y = runif(1000, 0, 1)

gfun = function(x)
  return(sqrt(1 - x ^ 2))

f1fun = function(x)
  return(x ^ 2)

f2fun = function(x)
  return(x)

gx = gfun(y)

f1x = f1fun(y)

b1 = lm(gx ~ f1x)$coef[2]

c(cor(gx,f1x),b1)

f2x = f2fun(y)

b2 = lm(gx ~ f2x)$coef[2]

c(cor(gx,f2x),b2)

theta_1 =
  b1 * (1 / 3) + gx - b1 * f1x

theta_2 =
  b2 * (1 / 2) + gx - b2 * f2x

result =
  tibble(
    model = c("gx", "U^2", "U"),
    theta = c(mean(gx), mean(theta_1), mean(theta_2)),
    effic = c(0, (var(gx) - var(theta_1)), (var(gx) - var(theta_2))) / var(gx)
  )

knitr::kable(result)
```


This result shows that $U^2$ has higher efficiency than $U$, the plot shows that,
$1 - U^2$ is more resemble to the original function. Corelation between simulated
 data shows that same conclusion(`r c(cor(gx,f1x),cor(gx,f2x))`)

```{r}
plot.function(gfun,0,1)
plot.function(function(x) 1-f1fun(x),0,1,add=T,col = "red")
plot.function(function(x) 1-f2fun(x),0,1,add=T,col = "blue")
```



# Problem 3
Obtain a Monte Carlo estimate of
$$\int_1^\infty \frac{x^2}{\sqrt{2\pi}} e^{-\frac{x^2}{2}} dx$$
by importance sampling and evaluate its variance. Write a  \em{\bf{R}} function to implement your procedure.

# Answer:  your answer starts here...

### Method 1
\begin{eqnarray}
\theta = \int{\frac{\frac{x^2}{\sqrt{2\pi}} e^{-\frac{x^2}{2}}*(I>1)+0*(I<=0)}
{Gamma(1,1/2)} *Gamma(1,1/2)}dx
\end{eqnarray}

```{r}
set.seed(123123)

y = rgamma(1000,1,1/2)

gfun = function(x)
  return(x ^ 2 / sqrt(2 * pi) * exp(-x ^ 2 / 2) * (x > 1))

gx = gfun(y)

y = dgamma(y,1,1/2)

theta = mean(gx/y)

theta
```

### Method 2 
We first us change of variables, where we take x = tan(y), s.t
\begin{eqnarray*}
\theta &=& \int_1^\infty \frac{x^2}{\sqrt{2\pi}} e^{-\frac{x^2}{2}}dx \\
&=& \int_{\pi/4}^{\pi/2}\frac{tan(y)^2}{\sqrt{2\pi}}* e^{-\frac{tan(y)^2}{2}}\frac{1}{cos(y)^2}dy\\
&=& \int_1^\infty \frac{x^2}{\sqrt{2\pi}} e^{-\frac{x^2}{2}}dx \\
&=& \int_{\frac{{\pi/4}^{\pi/2}\frac{tan(y)^2}{\sqrt{2\pi}}* e^{-\frac{tan(y)^2}{2}}\frac{1}{cos(y)^2}}{\frac{1}{(\pi/2-\pi/4)}}*\frac{1}{(\pi/2-\pi/4)}}dy
\end{eqnarray*}

```{r}
set.seed(123123)

y = runif(1000,pi/4,pi/2)

yfun =
  function(x) {
    return((tan(x)^2*exp(-tan(x)^2/2))/(sqrt(2*pi)*cos(x)^2))
  }

gx = yfun(y)

yx = 1/(pi/2 - pi/4)

theta = mean(gx/yx)

theta
```


### Method 3

$$\int_1^\infty x^2*\frac{1}{\sqrt{2\pi}} e^{-\frac{x^2}{2}} dx$$,

if we use $g(x) = x^2*I(x>1) + 0*I(x\le0)$ and introduce standard normal as p(x), the function becomes:
$$\int_1^\infty \frac{x^2*\frac{1}{\sqrt{2\pi}} e^{-\frac{x^2}{2}}}{\frac{1}{\sqrt{2\pi}} e^{-\frac{x^2}{2}}}\frac{1}{\sqrt{2\pi}} e^{-\frac{x^2}{2}} dx=E[x^2*I(x>1)]$$,
```{r}
set.seed(123123)

y = rnorm(1000)

gfun = function(x)
  return(x ^ 2* (x > 1))

gx = gfun(y)

theta = mean(gx)

theta
```
